---
title: "Exercise 4 - Relationship Between Examiner Centrality and Application Processing Time"
author: "Sheida Majidi"
date: "2024-04-06"
output:
  word_document: default
  html_document: default
---

# Introduction

In this exercise, I want to present my analysis on the relationship between patent examiner centrality within the advice network and the processing time of patent applications. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(gender)
library(wru)
library(igraph)
library(ggraph)
library(gridExtra)


```

# Data loading and preprocessing

First, I load the data and prepare it for the analysis:


```{r data}
applications = read_csv("/Users/sheidamajidi/Desktop/Winter2024/Winter2024-2/ORGB672/Exercises/Exercise 4/app_data_starter.csv", show_col_types = FALSE)
edges = read_csv("/Users/sheidamajidi/Desktop/Winter2024/Winter2024-2/ORGB672/Exercises/Exercise 4/edges_sample.csv",show_col_types = FALSE)

```

# Data Preprocessing

## Estimating examiner demographics


```{r examiner-demographic-estimation, echo=FALSE}

# Gender estimation

examiner_names=applications %>% distinct(examiner_name_first)

# Predict gender based on first names
examiner_names_gender=examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(examiner_name_first = name, gender, proportion_female)

# Join gender data back to the main applications dataset
applications=applications %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")

# Race estimation

examiner_surnames=applications %>% select(surname = examiner_name_last) %>% distinct()

# Predict race based on last names
examiner_race=predict_race(voter.file = examiner_surnames, surname.only = T) %>% as_tibble()

# Select the race with the highest probability for each last name
examiner_race=examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))

# Join race data back to the main applications dataset
applications=applications %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))


# Tenure calculation

examiner_dates=applications %>% 
  select(examiner_id, filing_date, appl_status_date)

# Convert dates to a consistent format
examiner_dates=examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))

# Calculate the earliest and latest dates for each examiner and their tenure in days
examiner_dates=examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
  ) %>% 
  filter(year(latest_date)<2018)

# Join tenure data back to the main applications dataset
applications=applications %>% left_join(examiner_dates, by = "examiner_id")


```

## Creating processing time variable

```{r processing-time-variable, echo=FALSE}
applications=applications %>%
  mutate(
    final_decision_date = coalesce(patent_issue_date, abandon_date),
    app_proc_time = as.numeric(difftime(final_decision_date, filing_date, units = "days"))
  )

```


# Centrality measures

First, I want to create a unique list of examiner ids before the analysis:

```{r unique-examiner-ids, echo=FALSE}
unique_examiner_ids=unique(c(edges$ego_examiner_id, edges$alter_examiner_id))

g=graph_from_data_frame(edges[, c("ego_examiner_id", "alter_examiner_id")], directed = TRUE, vertices = data.frame(name = unique_examiner_ids))
```

Now, it's time to proceed with calculating the centrality measures:

```{r centrality-measures, echo=FALSE}
centrality_entire=data.frame(
  examiner_id = V(g)$name,
  degree_centrality = degree(g, mode = "out"),
  betweenness_centrality = betweenness(g, directed = TRUE),
  closeness_centrality = closeness(g, mode = "out")
)

centrality_entire$examiner_id=as.numeric(centrality_entire$examiner_id)

applications=applications %>%
  left_join(centrality_entire, by = "examiner_id")

centrality_entire=data.frame(
  examiner_id = V(g)$name,
  degree_centrality = degree(g, mode = "out"),
  betweenness_centrality = betweenness(g, directed = TRUE),
  closeness_centrality = closeness(g, mode = "out")
)

centrality_entire$examiner_id=as.numeric(centrality_entire$examiner_id)

applications=applications %>%
  left_join(centrality_entire, by = "examiner_id")

```


# Exploratory Data Analysis
```{r column-names, echo=FALSE}
# column names to confirm the presence of 'gender'
colnames(applications)
```


```{r eda, echo=FALSE}
# Histogram of processing time
ggplot(applications, aes(x = app_proc_time)) +
  geom_histogram(binwidth = 30, fill = "blue", color = "black") +
  labs(title = "Histogram of Application Processing Time", x = "Processing Time (days)")

```

# Regression Analysis

First, I will remove the missing values in degree, betweenness, and closeness centrality.
```{r missing-value-remove, echo=FALSE}
applications_clean=applications %>%
  filter(!is.na(degree_centrality.x),
         !is.na(betweenness_centrality.x),
         !is.na(closeness_centrality.x))

```

## Degree centrality linear regression model

First, I do the analysis to estimate the linear regression model with degree_centrality as the independent variable

```{r regression-analysis-degree-centrality, echo=FALSE}
# Estimate the linear regression model with degree_centrality as the independent variable
degree_model=lm(
  app_proc_time ~ degree_centrality.x + gender.x + race.x + tenure_days.x,
  data = applications_clean
)

summary(degree_model)

```
### Explanation on degree centrality linear regression model

The degree_model includes degree_centrality, gender, race, and tenure_days as independent variables, and app_proc_time as the dependent variable. The adjusted R-squared value of the model is 0.003339, which means that only about 0.33% of the variation in app_proc_time can be explained by the model. This is quite low, indicating that the model does not fit the data well.

## Betweenness centrality linear regression model

Now, I do the analysis to estimate the linear regression model with betweenness_centrality as the independent variable

```{r regression-analysis-betweenness-centrality, echo=FALSE}
# Betweenness centrality linear regression model
betweenness_model=lm(
  app_proc_time ~ betweenness_centrality.x + gender.x + race.x + tenure_days.x,
  data = applications_clean
)
summary(betweenness_model)

```

### Explanation on betweenness centrality linear regression model

The betweenness_model includes betweenness_centrality, gender, race, and tenure_days as independent variables, and app_proc_time as the dependent variable. The adjusted R-squared value of the model is 0.003478, which is still low, suggesting that betweenness_centrality is not a good predictor of app_proc_time.

## Closeness centrality linear regression model

Next, I do the analysis to estimate the linear regression model with closeness_centrality as the independent variable

```{r closeness_model}
# Closeness centrality linear regression model
closeness_model=lm(
  app_proc_time ~ closeness_centrality.x + gender.x + race.x + tenure_days.x,
  data = applications_clean
)
summary(closeness_model)
```


### Explanation on closeness centrality linear regression model
The closeness_model includes closeness_centrality, gender, race, and tenure_days as independent variables, and app_proc_time as the dependent variable. The adjusted R-squared value of the model is 0.008616, which is slightly better than that of the betweenness_model, but still relatively low. This suggests that while closeness_centrality may have some predictive power for app_proc_time, it is not a strong predictor on its own.


## Combined model of linear regression

Lastly, I do the analysis to estimate the linear regression combined model

```{r combined_model, echo=FALSE}
# Combined centrality linear regression model
combined_model=lm(
  app_proc_time ~ degree_centrality.x + betweenness_centrality.x + closeness_centrality.x + gender.x + race.x + tenure_days.x,
  data = applications_clean
)
summary(combined_model)
```

### Explanation on combined model

The combined model (including degree, betweenness, and closeness centralities) has an adjusted R-squared of 0.008712, while the closeness_model has an adjusted R-squared of 0.008616 Although the combined model has a slightly higher adjusted R-squared, the improvement is not that significant


# Analysis to see if this relationship differ by examiner gender

## Degree-Gender interaction

```{r degree_gender_interaction}
# Degree centrality model with interaction
degree_gender_interaction=lm(
  app_proc_time ~ degree_centrality.x * gender.x + race.x + tenure_days.x,
  data = applications_clean
)
summary(degree_gender_interaction)
```

### Explanation on Degree-Gender interaction

In the degree-gender interaction model, there's a significant interaction between degree centrality and gender, indicating that the impact of degree centrality on processing time differs between male and female examiners. While degree centrality slightly increases processing time, its impact is lessened for male examiners, as suggested by the negative interaction term. However, the overall explanatory power remains low.


## Betweenness-Gender interaction

```{r betweenness_gender_interaction}
# Betweenness centrality model with interaction
betweenness_gender_interaction=lm(
  app_proc_time ~ betweenness_centrality.x * gender.x + race.x + tenure_days.x,
  data = applications_clean
)
summary(betweenness_gender_interaction)
```
### Explanation on Betweenness-Gender interaction

The betweenness-gender interaction model reveals a negative effect of betweenness centrality on processing time, which becomes more pronounced for male examiners (positive interaction term). This suggests that higher betweenness centrality could lead to longer processing times, especially for males. Still, the model explains a very small portion of the variability in processing time.



## Closeness-Gender interaction:

```{r closeness_gender_interaction}
# Closeness centrality model with interaction
closeness_gender_interaction=lm(
  app_proc_time ~ closeness_centrality.x * gender.x + race.x + tenure_days.x,
  data = applications_clean
)
summary(closeness_gender_interaction)
```

### Explanation on Closeness-Gender interaction

In the closeness-gender interaction model, closeness centrality shows a substantial negative impact on processing time, which is altered for male examiners, indicated by the negative interaction term. This implies that while closeness generally reduces processing time, this effect is less significant for male examiners.


## Combined-Gender interaction:

```{r combined_gender_interaction}
# Combined model with interaction
combined_gender_interaction=lm(
  app_proc_time ~ (degree_centrality.x + betweenness_centrality.x + closeness_centrality.x) * gender.x + race.x + tenure_days.x,
  data = applications_clean
)
summary(combined_gender_interaction)
```

### Explanation on Combined-Gender interaction

The combined-gender interaction model integrates all three centrality measures and their interactions with gender. It shows varied effects of centrality on processing time, with the interactions indicating that gender moderates these effects. Despite incorporating more factors, the model's adjusted R-squared remains low, indicating limited overall explanatory power.


# Conclusion

Across all models, the interaction terms highlight that gender plays a role in how centrality measures affect processing time. However, the low adjusted R-squared values in each model suggest that centrality measures and gender, while statistically significant in some cases, do not robustly predict application processing times. The consistent low explanatory power across models indicates that additional factors not captured here are likely influencing processing times at the USPTO. These findings underscore the complexity of the factors affecting processing times and suggest that a more nuanced model, possibly including more variables or different types of analysis, would be necessary to fully understand these dynamics.


